---
title: "Lab 1_2"
output: html_document
---

```{r}
getwd()
```

## Check to make sure you are up to date with current version of R and R Studio.
## Use option + command + i to add new chunk.
```{r}
R.Version()
```

## Check the current version of the specified package.
```{r}
packageVersion("tidyverse")
```

## Provides overview of the session.
```{r}
sessionInfo()
```

## Use "?" to pull up a help document.
```{r}
?mean
```

## R can be used as a calculator.
```{r}
2*3
```

## R can perform functions.
```{r}
mean(2, 3, 4, 5)
```

```{r}
cos(30)
```

## R's strength lies in objects.
```{r}
x <- 20
```

```{r}
print(x)
```

```{r}
x*2
```

```{r}
treatment<-36
control<-38

my_experiment<-treatment+control

my_experiment
```

```{r}
my_experiment*2
```

## Create different objects.
```{r}
my_numeric <- 42
my_integer <- 2L
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```

## Find out the class of different objects.
```{r}
class(my_numeric)
class(my_integer)
class(my_character)
class(my_logical)
class(my_complex)
```

