---
title: "Lab2_2_rev_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
getwd()
```

```{r}
fish <- readr::read_csv("~/Desktop/class_files-master/Gaeta_etal_CLC_data.csv")
```

#Load the Tidyverse
```{r}
library("tidyverse")
```

# Summmary Functions
##### choose the suite of summary functions you like the most and find the most helpful
```{r}
names(fish)
```

```{r}
head(fish)
```

```{r}
summary(fish)
```

```{r}
glimpse(fish)
```

```{r}
str(fish)
```

# Practice
##### 1. Load the data `mammal_lifehistories_v2.csv` and place it into a new object called `mammals`.

```{r}
mammals <- readr::read_csv("~/Desktop/class_files-master/mammal_lifehistories_v2.csv")
```

##### 2. Provide the dimensions of the data frame.

```{r}
dim(fish)
```

##### 3. Display the column names in the data frame. 

```{r}
names(fish)
```

##### 4. Use str() to show the structure of the data frame and its individual columns; compare this to glimpse(). 

```{r}
str(fish)
```

```{r}
glimpse(fish)
```


##### 5. Print out the first few rows of the data using the function head(). 

```{r}
head(fish)
```

# Filter
##### dplyr is part of the tidyverse
##### prof uses it almost everyday
##### first command to learn is filter
##### filter looks through dataframe and pulls out rows that fit criteria that you specify

```{r}
names(fish)
```

```{r}
filter(fish, lakeid=="AL")
```

##### can filter work for columns as well as rows?
```{r}
filter(fish, length>350)
```

```{r}
filter(fish, length==167 & length==175)
```
##### there are no fish that have both length 167 and length 175
##### want to use "or", instead of "and"
```{r}
filter(fish, length==167 | length==175)
```

##### "!" means "not"
##### so we are essentially saying, "filter fish for the lengths that are not 175"
```{r}
filter(fish, !length==175)
```

# Practice
##### 1. Filter the `fish` data to include the samples from lake `DY`.

```{r}
filter(fish, lakeid=="DY")
```

##### 2. Filter the data to include all lakes except AL.

```{r}
filter(fish, !lakeid=="AL")
```

##### 3. Filter the data to include all lakes except AL and DY.

```{r}
filter(fish, !lakeid=="AL" & !lakeid=="DY")
```

##### 4. Filter the data to include all fish with a scale length greater than or equal to 11.

```{r}
filter(fish, length>=11)
```

##### 5. Filter the data to include fish only from lake AL and with a scalelength greater than or equal to 2 and less than or equal to 4.

```{r}
filter(fish, lakeid=="AL", scalelength>=2 & scalelength<=4)
```

# Pipes
##### pipes make code cleaner
##### this essentially says, "take fish and pipe it to the function filter"
```{r}
fish %>% 
  filter(lakeid=="AL")
```

##### use select( )
##### if you don't really need to do anything with the data????
```{r}
fish %>%
  select(lakeid, scalelength)
```

##### can combine select with filter through the pipe
```{r}
fish %>%
  select(lakeid, scalelength) %>%
  filter(lakeid=="AL",
         scalelength>=2 & scalelength<=4)
```

##### to make it a dataframe, just have to assign it to new object
```{r}
fish_new <- fish %>%
  select(lakeid, scalelength) %>%
  filter(lakeid=="AL",
         scalelength>=2 & scalelength<=4)
```

### pipe -> select -> filter