---
title: "lab2_hw_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
getwd()
```

```{r}
library("tidyverse")
```

# 1
#### This data is taken from a PNAS publication by V.M. Savage and G.B. West.
```{r}
?msleep
```

# 2
#### Msleep has 11 columns detailing the sleep habits of 83 mammals, along with information about genus, "vore," order, and conservation status.
```{r}
names(msleep)
```
```{r}
glimpse(msleep)
```
```{r}
head(msleep)
```

# 3
```{r}
new_dataframe <- subset(msleep, select=c("name", "genus", "bodywt"))
```
```{r}
new_dataframe <- arrange(new_dataframe, desc(bodywt))
```
```{r}
print(new_dataframe)
```

# 4
```{r}
small_dataframe <- msleep %>%
  select(sleep_total, bodywt) %>%
  filter(bodywt<=200)
```
```{r}
print(small_dataframe)
```
```{r}
large_dataframe <- msleep %>%
  select(sleep_total, bodywt) %>%
  filter(bodywt>=200)
```
```{r}
print(large_dataframe)
```


# 5
#### The average sleep duration for large mammals is 3.3
```{r}
mean(large_dataframe$sleep_total)
```

# 6
#### The average sleep duration for small mammals is 11.09079.
```{r}
mean(small_dataframe$sleep_total)
```

# 7
#### The little brown bat, big brown bat, thick-tailed opposum, giant armadillo, and North American Opposum sleep 18 hours or more.
```{r}
eighteen_hours <- msleep %>%
   select(name,genus, order, sleep_total) %>%
   filter(sleep_total>=18)
eighteen_hours <- arrange(eighteen_hours,desc(sleep_total))
```
```{r}
print(eighteen_hours)
```

