---
title: "Lab5_1_rev_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(skimr)
```

# Grammar of Graphics
### Plots built in layers

### Make plots visibly appealing, but not too confusing
### discrete data vs. continuous data vs. categorical data
### plot you makes depends on what kind of data you have in your data set

### What is our data set?
```{r}
iris <- iris
```

### data is iris, mapping is aesthetic
### set y axis and x axis
### specifying axis is specifying a column
## geom is box plot??

```{r}
ggplot(data=iris, mapping=aes(x=Species, y=Petal.Length))+
  geom_boxplot()
```

# Practice
### 1. Take a moment to practice. Use the iris data to build a scatterplot that compares sepal length vs. sepal width. Use the cheatsheet to find the correct `geom_` for a scatterplot.

```{r}
ggplot(data=iris, mapping=aes(x=Sepal.Length, y=Sepal.Width))+
  geom_point()
```

### Load the data
```{r}
homerange <- readr::read_csv("/Users/sophieborison/Desktop/class_files-master/Tamburelloetal_HomeRangeDatabase.csv")
```

```{r}
homerange <- 
  homerange %>%  
  select(common.name, taxon, family, genus, species, mean.mass.g, log10.mass, mean.hra.m2, log10.hra, preymass, log10.preymass, trophic.guild)
```

```{r}
names(homerange)
```

### Is there a correlation between organism size and the size of its habitat?
### geom jitter helps with noise
### can add regression lign
```{r}
ggplot(data=homerange, mapping=aes(x=log10.mass, y=log10.hra)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE)
```

### R already knows mapping
### Our data already implies y-axis, it is number of observations
### Can't specify y for a geom bar, it is just count
### Especially helpful if you want to make counts of catergories
```{r}
homerange %>%
  ggplot(aes(x=trophic.guild)) +
  geom_bar(stat="count")
```

```{r}
carni_mammals <- 
  homerange %>% 
  filter(taxon=="mammals", 
         trophic.guild=="carnivore") %>% 
  group_by(family) %>% 
  summarize(count=n(),
            mean_body_wt=mean(log10.mass)) %>% 
  arrange(desc(mean_body_wt))
carni_mammals
```

### mean bodyweight by ___ of carnivorous mammals
### need to tell it not to do counts, there is nothing to count
```{r}
carni_mammals%>% 
  ggplot(aes(x=family, y=mean_body_wt))+
  geom_bar(stat="identity")
```

### theme controls text aspects, etc.
### angle controls text angle
### hjust sets text spacing
### one of those axes needs to be a categorical variable
# I'm confused what's going on, what does theme, angle, etc. do?
```{r}
carni_mammals%>% 
  ggplot(aes(x=family, y=mean_body_wt))+
  geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 60, hjust=1))
```

### reorder makes it descending
### coord_flip does ___
```{r}
carni_mammals%>% 
  ggplot(aes(x=reorder(family, -mean_body_wt), y=mean_body_wt))+
  geom_bar(stat="identity")+
  coord_flip()
```

# Practice

### Filter the `homerange` data to include `mammals` only.
```{r}
mammals <- 
  homerange %>% 
  filter(taxon=="mammals") %>% 
  select(common.name, family, genus, species, trophic.guild, mean.mass.g, log10.mass, mean.hra.m2, log10.hra, preymass, log10.preymass)
```

```{r}
mammals
```

### 1. Are there more herbivores or carnivores in mammals? Make a bar plot that shows their relative proportions.

# Identity vs. count?? I'm confused.
```{r}
mammals %>% 
  ggplot(aes(x=trophic.guild))+
  geom_bar(stat="count")+
  theme(axis.text.x = element_text(angle = 60, hjust=1))
```


### 2. Is there a positive or negative relationship between mass and homerange? How does this compare between herbivores and carnivores? Make two plots that illustrate these eamples below.  

```{r}
mammals %>%
  filter(trophic.guild=="herbivore") %>%
  ggplot(mapping=aes(x=log10.mass, y=log10.hra)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE)
```

```{r}
mammals %>%
  filter(trophic.guild=="carnivore") %>%
  ggplot(mapping=aes(x=log10.mass, y=log10.hra)) +
  geom_point() +
  geom_smooth(method=lm, se=FALSE)
```

### 3. Make a barplot that shows the masses of the top 10 smallest mammals in terms of mass. Be sure to use `stat'="identity"`.

```{r}
mammals %>% 
  top_n(-10, log10.mass) %>% 
  ggplot(aes(x=common.name, y=log10.mass))+
  geom_bar(stat="identity")+
  coord_flip()
```

### skip 8,9,10 on homework

### can store plots as objects and layer them
```{r}
qqq <- mammals %>%
  filter(trophic.guild=="herbivore")
```

```{r}
p1 <- ggplot(data=mammals, mapping=aes(x=log10.mass, y=log10.hra))
```

```{r}
p1+geom_point()+geom_smooth(method=lm, se=FALSE)
```

```{r}
p1+geom_line()+geom_point()
```
