---
title: "lab5_hw_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
```

## Mammals Life History
Let's revisit the mammal life history data to practice our `ggplot` skills. Some of the tidy steps will be a repeat from the homework, but it is good practice. The [data](http://esapubs.org/archive/ecol/E084/093/) are from: *S. K. Morgan Ernest. 2003. Life history characteristics of placental non-volant mammals. Ecology 84:3402.*

# 1. Load the data.
```{r}
lifehistories <- readr::read_csv("/Users/sophieborison/Desktop/class_files-master/data/mammal_lifehistories_v2.csv")
```

# 2. Use your preferred function to have a look. Do you notice any problems?

### There are many NAs.
```{r}
skimr::skim(lifehistories)
```

# 3. There are NA's. How are you going to deal with them?

### Repalace -999 with "NA"
```{r}
lifehistories_NA <- 
  lifehistories %>% 
  na_if(-999.00)
lifehistories_NA
```

# 4. Where are the NA's? This is important to keep in mind as you build plots.

### Most of the NA's are in wean mass, max. life, and litters/year.
```{r}
lifehistories_NA %>%
  purrr::map_df(~ sum(is.na(.))) %>%
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```

# 5. Some of the variable names will be problematic. Let's rename them here as a final tidy step.

```{r}
lifehistories_NA %>%
  mutate(
          genus        = Genus,
          wean_mass    = `wean mass`,
          max_life     = `max. life`,
          litter_size  = `litter size`,
          litters_yr   = `litters/year`
          )
```



##`ggplot()`
For the questions below, try to use the aesthetics you have learned to make visually appealing and informative plots. Make sure to include labels for the axes and titles.
```{r}
options(scipen=999) #cancels the use of scientific notation for the session
```

# 6. What is the relationship between newborn body mass and gestation? Make a scatterplot that shows this relationship. 

```{r}
ggplot(data=lifehistories_NA, mapping=aes(x=newborn, y=gestation)) +
  geom_point()
  #geom_smooth(method=lm, se=FALSE)
```

# 7. You should notice that because of the outliers in newborn mass, we need to make some changes. We didn't talk about this in lab, but you can use `scale_x_log10()` as a layer to correct for this issue. This will log transform the y-axis values.

```{r}
ggplot(data=lifehistories_NA, mapping=aes(x=newborn, y=gestation)) +
  geom_point() +
  scale_x_log10()
```

# 8. Now that you have the basic plot, color the points by taxonomic order.

```{r}
ggplot(data=lifehistories_NA, mapping=aes(x=newborn, y=gestation)) +
  geom_point(aes(colour=order)) +
  scale_x_log10()
```

# 9. Lastly, make the size of the points proportional to body mass.

```{r}
ggplot(data=lifehistories_NA, mapping=aes(x=newborn, y=gestation)) +
  geom_point(aes(colour=order, size=mass)) +
  scale_x_log10()
```

# 10. Make a plot that shows the range of lifespan by order.

### n/a

