---
title: "lab6_hw_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library("RColorBrewer")
```

# Install gapminder
```{r}
install.packages("gapminder")
```

```{r}
library("gapminder")
```

# Load the data into a new object called gapminder
```{r}
gapminder <- 
  gapminder::gapminder
```

# 1
### Explore the data using the various function you have learned. Is it tidy, are there any NA's, what are its dimensions, what are the column names, etc.

```{r}
gapminder %>%
  purrr::map_df(~ sum(is.na(.))) %>%
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```

```{r}
names(gapminder)
```

```{r}
glimpse(gapminder)
```

# 2
### We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer on average. Make a quick plot below to visualize this relationship.

```{r}
gapminder %>%
  ggplot(aes(x=gdpPercap, y=lifeExp)) +
  geom_point()
```

# 3
### There is extreme disparity in per capita GDP. Rescale the x axis to make this easier to interpret. How would you characterize the relationship?

### This is a positive correlation.
```{r}
gapminder %>%
  ggplot(aes(x=gdpPercap, y=lifeExp)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method=lm, se=FALSE)
```

# 4
### This should look pretty dense to you with significant overplotting. Try using a faceting approach to break this relationship down by year.

```{r}
gapminder %>%
  ggplot(aes(x=gdpPercap, y=lifeExp)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method=lm, se=FALSE) +
  facet_wrap(~year)
```

# 5
### Simplify the comparison by comparing only 1952 and 2007. Can you come to any conclusions?

```{r}
gapminder %>%
  filter(year == 1952 | year == 2007) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method=lm, se=FALSE) +
  facet_wrap(~year)
```

# 6
### Let's stick with the 1952 and 2007 comparison but make some aesthetic adjustments. First try to color by continent and adjust the size of the points by population. Add `+ scale_size(range = c(0.1, 10), guide = "none")` as a layer to clean things up a bit.

```{r}
gapminder %>%
  filter(year == 1952 | year == 2007) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop)) +
  scale_size(range = c(0.1, 10), guide = "none") +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method=lm, se=FALSE) +
  facet_wrap(~year)
```

# 7
### Although we did not introduce them in lab, ggplot has a number of built-in themes that make things easier. I like the light theme for these data, but you can see lots of options. Apply one of these to your plot above.

```{r}
?theme_light
```

```{r}
gapminder %>%
  filter(year == 1952 | year == 2007) %>%
  ggplot(aes(x=gdpPercap, y=lifeExp, color=continent, size=pop)) +
  scale_size(range = c(0.1, 10), guide = "none") +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method=lm, se=FALSE) +
  facet_wrap(~year) +
  theme_light()
```

# 8
### What is the population for all countries on the Asian continent in 2007? Show this as a barplot.

```{r}
gapminder %>%
  filter(year == 2007, continent=="Asia") %>%
  ggplot(aes(x=reorder(country, pop), y=pop)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle=90))
```

#9
### You should see that China's population is the largest with India a close second. Let's focus on China only. Make a plot that shows how population has changed over the years.

```{r}
gapminder %>%
  filter(country=="China") %>%
  ggplot(aes(x=year, y=pop)) +
  geom_point()
```

# 10
### Let's compare China and India. Make a barplot that shows population growth by year using `position=dodge`. Apply a custom color theme using RColorBrewer.

```{r}
gapminder %>%
  filter(country=="China" | country == "India") %>%
  ggplot(aes(x=year, y=pop, group=country, fill=country)) +
  geom_bar(stat="identity", position="dodge") +
  scale_fill_brewer(palette = "Accent")
```