---
title: "lab3_hw_SBUCD"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries
```{r}
library(tidyverse)
```

```{r}
fisheries <- readr::read_csv(file = "/Users/sophieborison/Desktop/class_files-master/data/FAO_1950to2012_111914.csv")
```

# 1
### Do you see any potential problems with the column names? Does the error message now make more sense?  

##### The column names are duplicates. R "deduplicated" them.

# 2
### The `make.names()` command is helpful when there are issues with column names. Notice that although the names are still cumbersome, much of the problemtatic syntax is removed.

##### Make names.
```{r}
names(fisheries) = make.names(names(fisheries), unique=T)
names(fisheries)
```

# 3
###  Let's rename the columns. Use `rename()` to adjust the names as follows. Double check to make sure the rename worked correctly. Make sure to replace the old fisheries object with a new one so you can keep the column names.

##### Rename columns using a pipe. 
```{r}
renamed_fisheries <-
  fisheries %>%
  dplyr::rename(
    country = Country..Country.,
    commname = Species..ASFIS.species.,
    sciname = Species..ASFIS.species._2,
    spcode = Species..ASFIS.species._1,
    spgroup = Species..ISSCAAP.group.,
    spgroupname = Species..ISSCAAP.group._1,
    region = Fishing.area..FAO.major.fishing.area.,
    unit = Measure..Measure.
  )
```
```{r}
names(renamed_fisheries)
```

#4
### Are these data tidy? Why or why not, and, how do you know?

##### These data are not yet tidy because instead of "year" being a column name, the years themselves are column names.
```{r}
print(renamed_fisheries)
```

# 5
###  We need to tidy the data using `gather()`. The code below will not run because it is commented (#) out. I have added a bit of code that will prevent you from needing to type in each year from 1950-2012 but you need to complete the remainder `QQQ` and remove the `#`.

##### Use gather to tidy the data.
```{r}
fisheries_tidy <- 
  renamed_fisheries %>% 
  gather(num_range('X',1950:2012), key='year', value='catch')
```

# 6
### Use `glimpse()` to look at the categories of the variables. Pay particular attention to `year` and `catch`. What do you notice?

##### There are a lot of missing data in `catch`.
```{r}
glimpse(fisheries_tidy)
```

#7
### From question 6 you should see that there are a lot of entries that are missing. In R, these are referred to as NA's but they can be coded in different ways by the scientists in a given study. In order to make the data tidy, we need to deal with them. As a preview to our next lab, run the following code by removing the `#`. It removes the 'X' from the years and changes the `catch` column from a character into a numeric. This forces the blank entries to become NAs. The error "NAs introduced by coercion" indicates their replacement.

```{r}
fisheries_tidy <- 
  fisheries_tidy %>% 
  mutate(
    year= as.numeric(str_replace(year, 'X', '')), #remove the X from year
    catch= as.numeric(str_replace(catch, c(' F','...','-'), replacement = '')) #replace character strings with NA
    )
```

# 8
### Are the data tidy? Why?  

##### The data are tidy because there is one observation per row, and each column name is a category rather than a variable.

# 9
###  You are a fisheries scientist studying cephalopod catch during 2008-2012. Identify the top five consumers (by country) of cephalopods (don't worry about species for now). Restrict the data frame only to our variables of interest.

##### The top five consumers of cephalopods are India, China, Madagascar, Mozambique, and Taiwan.
```{r}
fisheries_tidy %>% 
  select(country, sciname, year, catch) %>% 
  filter(sciname=="Cephalopoda") %>%
  filter(year== 2012) %>%
  arrange(desc(catch))
```

# 10
### Let's be more specific. Who consumes the most `Common cuttlefish`? Store this as a new object `cuttle`.

##### France consumes the most Common cuttlefish.
```{r}
cuttle <-
  fisheries_tidy %>%
  filter(commname=="Common cuttlefish") %>%
  select(country, commname, catch) %>%
  arrange(desc(catch))
```
```{r}
print(cuttle)
```

