---
title: "lab4_hw_rev_SBUCD"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the tidyverse
```{r}
library(tidyverse)
```

# Load mammal data
```{r}
life_history <- readr::read_csv("/Users/sophieborison/Desktop/class_files-master/data/mammal_lifehistories_v2.csv")
```

# Rename some of the variables
```{r}
life_history <- 
  life_history %>% 
  dplyr::rename(
          genus        = Genus,
          wean_mass    = `wean mass`,
          max_life     = `max. life`,
          litter_size  = `litter size`,
          litters_yr   = `litters/year`
          )
```

# 1.
### Explore the data using the method that you prefer. Below, I am using a new package called `skimr`. If you want to use this, make sure that it is installed.

### Install `skimr`
```{r}
# install.packages("skimr")
```

### Load `skimr`
```{r}
library("skimr")
```

### Use `skimr`
```{r}
life_history %>% 
  skimr::skim()
```

# 2.
### Run the code below. Are there any NA's in the data? Does this seem likely?
```{r}
life_history %>% 
  summarize(number_nas= sum(is.na(life_history)))
```

### R reports that there are no NA's in the data. This is unlikely. 

# 3.
### Are there any missing data (NA's) represented by different values? How much and where? In which variables do we have the most missing data? Can you think of a reason why so much data are missing in this variable?

### There are 1440 NA's represented by -999. There are the most NA's in wean mass, max life, and litters year. These variables may be difficult to measure in a species, or the species may be difficult to find or hard to get to. What are the chances of finding a weaning endangered South Andean deer in the high elevation Andes?

### Count number of -999's
```{r}
life_history %>%
 count(-999)
```

### Replace -999 with NA
```{r}
life_history_na <- 
  life_history %>% 
  na_if(-999)
```

### Sort the NA's
```{r}
life_history_na %>% 
  purrr::map_df(~sum(is.na(.))) %>% 
  tidyr::gather(variables, num_nas) %>% 
  arrange(desc(num_nas))
```

# 4.
### Compared to the `msleep` data, we have better representation among taxa. Produce a summary that shows the number of observations by taxonomic order.

```{r}
life_history_na %>%
  group_by(order) %>%
  summarize(total=n())
```

# 5
### Mammals have a range of life histories, including lifespan. Produce a summary of lifespan in years by order. Be sure to include the minimum, maximum, mean, standard deviation, and total n.

```{r}
life_history_na %>%
  group_by(max_life) %>%
  summarize(min_max_life=min(max_life, na.rm=TRUE),
          max_max_life=max(max_life, na.rm=TRUE),
          mean_max_life=mean(max_life, na.rm=TRUE),
          stdev_max_life=sd(max_life, na.rm=TRUE),
          total=n())
```

# 6
### Let's look closely at gestation and newborns. Summarize the mean gestation, newborn mass, and weaning mass by order. Add a new column that shows mean gestation in years and sort in descending order. Which group has the longest mean gestation? What is the common name for these mammals?

### Proboscidea has the longest mean gestation.
```{r}
life_history_na %>%
  select(order, gestation, newborn, wean_mass) %>%
  group_by(order) %>%
  summarize(mean_gestation=mean(gestation, na.rm=TRUE),
            mean_newborn=mean(newborn, na.rm=TRUE),
            mean_wean_mass=mean(wean_mass, na.rm=TRUE)) %>%
  arrange(desc(mean_gestation))
```




