---
title: "lab3_1_rev_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load tidyverse
```{r}
library(tidyverse)
```

# review
### read the csv
```{r}
fish <- readr::read_csv("/Users/sophieborison/Desktop/class_files-master/Gaeta_etal_CLC_data.csv")
```

### filter
```{r}
glimpse(fish)
```
```{r}
names(fish)
```

```{r}
fish %>%
  filter(lakeid=="AL",
         length==167)
```

##### might be distracting to have all those columns
##### so use function select( ) to only show certain columns
```{r}
fish %>%
  select(fish_id, lakeid, scalelength)
```
### can use contains( ) function instead
### comes in handy if you have a data frame with lots of columns
```{r}
fish %>%
  select(contains("length"))
```

### can use ends_with( ) function as well
```{r}
fish %>%
  select(ends_with("mm"))
```


### oftentimes, we think that because we have no errors, everything is working
### couldve worked for the wrong reasons and not done what you thought it did

# Practice
###1. Build a dataframe focused on the lakes `AL` and `AR` and looking at radii length between 2 and 4 only. Use pipes!

```{r}
fish %>%
  filter(lakeid=="AL" | lakeid=="AR") %>%
  filter(radii_length_mm>=2 & radii_length_mm<=4)
```

### Arrange allows us to sort values in a column
```{r}
fish %>%
  arrange(desc(scalelength))
```

### Count the number of observations in ___
```{r}
fish %>%
  count(lakeid)
```

### mutate( ) allows use to create new columns in a data frame
```{r}
new_fish <- fish %>%
  select(lakeid, fish_id, scalelength, length) %>%
  filter(lakeid=="AL") %>%
  arrange(fish_id) %>%
  mutate(scale_ratio=(length/scalelength))
```

```{r}
names(new_fish)
```

# Practice
### 1. Build a dataframe focused on the scalelengths of `fish_id` 300 and 301. Use `arrange()` to sort from smallest to largest scalelength. Use pipes!

```{r}
fish %>%
  select(fish_id, scalelength) %>%
  filter(fish_id==300 | fish_id==301) %>%
  arrange(scalelength)
```

### 2. Add a new column to the fish data that is radii_length_mm divided by scalelength. Add another column that scales this number to a percentage
```{r}
new_new_fish <- fish %>%
  mutate(radii_scale=(100*(radii_length_mm/scalelength)))
```

```{r}
print(new_new_fish)
```

# can select, filter, and mutate a dataframe
# r makes things reproducable, which is important as data gets bigger and messier
