---
title: "lab4_1_rev_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the tidyverse
```{r}
library("tidyverse")
```

## tidy data includes that every column is its own vector

# Practice
```{r}
?USArrests
```

```{r}
arrests <- 
  USArrests %>% 
  mutate(State=rownames(USArrests)) %>% 
  select(State, Murder, Assault, Rape)
arrests
```

# 1. Are these data tidy? Please use `gather()` to tidy the data.

```{r}
tidy_arrests <-
  arrests %>%
  gather(Murder, Assault, Rape, key="Crime", value="per100k")
tidy_arrests
```


# 2. Restrict the data to assualt only. Sort in ascending order.

```{r}
tidy_arrests %>%
  filter(Crime=="Assault") %>%
  arrange(desc(per100k))
```

# Spread
### spread is the opposite of gather
### interchange between gathered data
### long format vs. wide format
### gathered = long format, spread = wide format
## important in order to make plots

### In the example below, `cases` and `population` represent variable names not observations.
```{r}
country <- c("Afghanistan", "Afghanistan", "Afghanistan", "Afghanistan", "Brazil", "Brazil", "Brazil", "Brazil", "China", "China", "China", "China")
year <- c("1999", "1999", "2000", "2000", "1999", "1999", "2000", "2000", "1999", "1999", "2000", "2000")
key <- c("cases", "population", "cases", "population", "cases", "population", "cases", "population", "cases", "population", "cases", "population")
value <- c(745, 19987071, 2666, 20595360, 37737, 172006362, 80488, 174504898, 212258, 1272915272, 213766, 1280428583)

tb_data <- data.frame(country=country, year=year, key=key, value=value)
tb_data
```

### When using `spread()` the `key` is the variable that you are spreading.
```{r}
tb_data %>% 
  spread(key=key, value=value)
```

# Practice
### 1. Run the following to build the `gene_exp` data frame.
```{r}
id <- c("gene1", "gene1", "gene2", "gene2", "gene3", "gene3")
type <- c("treatment", "control", "treatment", "control","treatment", "control")
L4_values <- rnorm(6, mean = 20, sd = 3)
```

```{r}
gene_exp <- 
  data.frame(gene_id=id, type=type, L4_values=L4_values)
gene_exp
```

### 2. Are these data tidy? Please use `spread()` to tidy the data.

### These data are not tidy because there are two different values in one column. (??)
```{r}
gene_exp %>%
  spread(type, L4_values)
```

# Summarize
# can summarize mean, min, max, median, etc.
```{r}
?msleep
```

From homework 2.
```{r}
large <- msleep %>% 
  select(name, genus, bodywt, sleep_total) %>% 
  filter(bodywt>=200) %>% 
  arrange(desc(bodywt))
```

```{r}
mean(large$sleep_total)
```

We can accomplish the same task using the `summarize()` function in the tidyverse and make things cleaner.
# make new column for mean of sleep total
```{r}
msleep %>% 
  filter(bodywt>=200) %>%
  summarize(mean_sleep_lg=mean(sleep_total))
```

You can also combine functions to make useful summaries for multiple variables.
```{r}
msleep %>% 
    filter(bodywt>=200) %>% 
    summarize(mean_sleep_lg = mean(sleep_total), 
              min_sleep_lg = min(sleep_total),
              max_sleep_lg = max(sleep_total),
              total = n())
```

# Practice

### 1. How many genera are represented in the msleep data frame?

```{r}
msleep %>%
  summarize(ngenera=n_distinct(genus))
```

### 2. What are the min, max, and mean body weight for all of the mammals? Be sure to include the total n.

```{r}
msleep %>%
  summarize(min = min(sleep_total),
            max = max(sleep_total),
            mean = mean(bodywt),
            total = n())
```

# group_by()
### group and summarize
### good way to see what a set of data are like
### take msleep and group it by "vore" type
```{r}
msleep %>%
  group_by(vore) %>% #we are grouping by feeding ecology
  summarize(min_bodywt=min(bodywt),
            max_bodywt=max(bodywt),
            mean_bodywt=mean(bodywt),
            total=n())
```

# Practice
### 1. Calculate mean brain weight by taxonomic order in the msleep data.

```{r}
msleep %>%
  group_by(order) %>%
  summarize(mean_brainwt=mean(brainwt))
```

### 2. What does `NA` mean?

### `NA` means that data is missing.

### 3. Try running the code again, but this time add `na.rm=TRUE`. What is the problem with Cetacea?

### Even if there is one `NA`, R can't calculate ___
### "remove those rows for which brain weight is NA"
```{r}
msleep %>%
  group_by(order) %>%
  summarize(mean_brainwt=mean(brainwt, na.rm=TRUE))
```
### Cetacea has NAs
### Can't calculate the mean of `NA`
```{r}
msleep %>% 
  filter(order=="Cetacea")
```
