---
title: "lab4_2_rev_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### NA's one of the big plights of data analysis
### in any field, in any science, there wil always be NA values
### even more common in big data sets
### how are those NA's represented? what did you mean by having that NA there?

### could have r summarize te number of NAs
### only things that R recognizes as NAs, but may not actually represent the number of NAs in the dataset
### doesnt tell you where NAs R

## Load the tidyverse
```{r}
library("tidyverse")
```

### `purrr` takes a command and repeats it across for you
### went through and told me where all my NAs were in the dataframe
```{r}
msleep_na <- 
  msleep %>%
  purrr::map_df(~ sum(is.na(.))) #map to a new data frame the sum results of the is.na function for all columns
msleep_na
```

### use gather to make it easier to read
```{r}
msleep %>% 
  purrr::map_df(~ sum(is.na(.))) %>% 
  tidyr::gather(key="variables", value="num_nas") %>% 
  arrange(desc(num_nas))
```

### count will also give you number of NAs in given column
```{r}
msleep %>% 
  count(conservation)
```

### can replace other notations of missing data with "NA"
Once you have an idea of how NA's are represented in the data, you can replace them with `NA` so that R can better deal with them. The bit of code below is very handy, especially if the data has NA's represented as **actual values that you want replaced** or if you want to make sure any blanks are treated as NA.
```{r}
msleep_na2 <- 
  msleep %>% 
  na_if("") #replace x data with NA
msleep_na2
```

# Practice
### 1. Did replacing blanks with NA have any effect on the msleep data? Demonstrate this using some code.

