---
title: "lab5_2_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
homerange <-
  readr::read_csv("/Users/sophieborison/Desktop/class_files-master/data/Tamburelloetal_HomeRangeDatabase.csv", na = c("", " ", "NA", "#N/A", "-999"))
```

```{r}
glimpse(homerange)
```

#Box Plots
### Box Plots are for range of values
```{r}
homerange %>% 
  ggplot(aes(x=class, y=log10.mass))+
  geom_boxplot()
```

# Practice

### 1. There are more herbivores than carnivores in the homerange data, but how do their masses compare? Make a boxplot that compares their masses. Use `log10.mass`.

```{r}
homerange %>%
  ggplot(aes(x=trophic.guild, y=log10.mass))+
  geom_boxplot()
```
### identity is for count, box plot is for range, so don't use identity

### 2. Practice your dplyr skills and filter out carnivorous mammals. 

### 3. Now use a boxplot to visualize the range of body mass by family of carnivore.

# Other Aesthetics

### Title is another layer, not piping anything
### Labs is for labels customized for your plot
```{r}
ggplot(data=homerange, mapping=aes(x=log10.mass, y=log10.hra)) +
  geom_point()+
         labs(title="Mass vs. Homerange",
              x = "Mass (log10)",
              y = "Homerange (log10)")
```

### Make type bold
### Change type size
```{r}
ggplot(data=homerange, mapping=aes(x=log10.mass, y=log10.hra)) +
  geom_point()+
  labs(title="Mass vs. Homerange",
              x = "Mass (log10)",
              y = "Homerange (log10)")+
  theme(plot.title=element_text(size=18, face="bold"),
        axis.text=element_text(size=12),
        axis.title=element_text(size=12))
```

### rel decides for you the size of titles relatively
```{r}
ggplot(data=homerange, mapping=aes(x=log10.mass, y=log10.hra)) +
  geom_point()+
  labs(title="Mass vs. Homerange",
              x = "Mass (log10)",
              y = "Homerange (log10)")+
  theme(plot.title = element_text(size = rel(1.5)))
```

# Practice

### 1. Make a barplot that shows counts of ectotherms and endotherms. Label the axes, provide a title, and fill by thermoregulation type.

```{r}
homerange %>%
  ggplot(aes(x=locomotion)) +
  geom_bar() +
  labs(title = "Locomotion vs. Individuals",
       x = "Locomotion Type",
       y = "Individuals") +
  theme(plot.title = element_text(size=14, face = "bold", hjust=0.5),
        axis.text = element_text(size=10),
        axis.title = element_text(size=12))
```

### 2. Make a boxplot that compares thermoregulation type by log10.mass. Group and fill by class. Label the axes and provide a title.

# Color

### Color points a certain color is different than coloring by a certain variable

```{r}
homerange %>%
  ggplot(aes(x=log10.mass, y=log10.hra))+
  geom_point(color.by(log10.mass))
```

```{r}
homerange %>%
  ggplot(aes(x=log10.mass, y=log10.hra))+
  geom_point(color="hotpink")
```

# Size

### Adjust the size of points

# Fill
```{r}
homerange %>%
  
```

# Group
```{r}
homerange %>%
  ggplot(aes(x=class, y=log10.mass, group=taxon, fill=taxon))+
  geom_boxplot()
```


