---
title: "labs_1_2_review_SBUCD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Arithmetic
### R is a calculator
```{r}
4^3
```

# Objects
### Objects can store data of multiple classes
```{r}
my_age <- 19
```
```{r}
my_age + 1
```

# Data Classes
### R works with multiple classes of data including integer, numeric, and character
```{r}
integer <- 3L
numeric <- 27.4
character <- "hello"
```
### Use the class( ) function to find out the class of an object
```{r}
class(integer)
class(numeric)
class(character)
```
### Logical tests of class type can also be performed
```{r}
is.integer(numeric())
```

# Vectors
### Vectors contain data of the same class
```{r}
happy_vector <- c(5, 10, 16, 17, 30)
happy_vector
```
```{r}
class(happy_vector)
```
### If there are different classes, R changes the content of the vector into what it thinks is the best class
```{r}
unhappy_vector <- c(5, 10, 16, "big", "small")
unhappy_vector
```
```{r}
class(unhappy_vector)
```

### Vectors may contain data of many different class types
```{r}
options(scipen=999) #disables scientific notation
organism<-c("Human","Mouse","Fruit Fly", "Roundworm","Yeast")
genomeSizeBP<-c(3000000000,3000000000,135600000,97000000,12100000)
estGeneCount<-c(30000,30000,13061,19099,6034)
```
### Combine them into a single vector
```{r}
genome_size <- c(genomeSizeBP, estGeneCount)
genome_size
```
### Provide names
```{r}
names(genome_size) <- c("genome size BP", "genome size BP", "genome size BP", "genome size BP", "genome size BP", "gene count", "gene count", "gene count", "gene count", "gene count")
genome_size
```
### You can select elements of a vector using `[]`
```{r}
humans <- genome_size[c(1, 6)]
humans
```

# Data Matrices
### Data matrices are slightly more efficient than vectors
```{r}
genome_matrix <- matrix(genome_size, nrow=2, byrow=T)
genome_matrix
```
### Name the columns and rows
```{r}
colnames(genome_matrix) <- organism
rownames(genome_matrix) <- c("genome size BP", "gene count")
genome_matrix
```
### Can use functions to make a calculation
```{r}
means <- rowMeans(genome_matrix)
```
### Add them to the matrix
```{r}
genome_matrix2 <- cbind(genome_matrix, means)
genome_matrix2
```

# Data Frames
### Data frames are like data matrices except every column is a vector
```{r}
genome_frame <- data.frame(organism=organism, genomeSizeBP=genomeSizeBP, estGeneCount=estGeneCount)
genome_frame
```

# Summary
### R works in vectors
### you can organisze those vectors in a data matrix
### but it is most efficient to use a data frame
### because we are working with multiple classes of data

###### need to make sure there are same number of names as there are values

#Next: tidyverse suite of packages